Estimating SARS-CoV-2 transmission in England from randomly sampled prevalence surveys

### 2023-11-21 incidence-prevalence.tex
correct some notation/terminology

Duration, or time-to-event, distributions can be expressed in several equivalent ways.
The duration distribution I will be concerned with is the distribution of the number of days for which an individual is prevalent.
Consider an individual $i$ infected on day $B_i$ that is prevalent during the interval $[B_i, E_i]$.
I define this period as an \emph{infection episode}, beginning on day $B_i$ and ending on day $E_i$.
$D_i = E_i - B_i + 1$ is their \emph{duration of positivity}, the number of days for which they are prevalent.
I model each $D_i$ as a discrete independent and identically distributed (iid) random variable; extending the methodology to account for covariates is left to further work.
The duration is a positive integer, \ie $D_i \in \{1, 2, \dots, \dmax \}$, where $\dmax$ is the maximum duration.
The first way of expressing the distribution of $D_i$ is as the standard probability mass function (pmf), $f_{D}(d) = \prob(D = d)$; or cumulative density function (CDF), $F_{D}(d) = \prob(D \leq d) = \sum_{i=1}^d f_{D}(i)$.
The \emph{survival function} gives the probability that the duration is at least $d$ days: $S(d) = \prob(D \geq d) = 1 - F_{D}(D - 1)$ with $S(1) = 1$ (a common alternative convention for survival analysis with discrete variables is to define the survival function as $\prob(D >d)$).
The \emph{hazard} on day $d$ is the probability of recovering at time $d$ conditional on having survived to time $d$: $\lambda(d) = \prob(D = d \mid D \geq d) = f_{D}(d) / S(d)$.
Throughout this thesis, the survival function will be the primary quantity of interest, and it will normally be parameterised in terms of the hazard; these two quantities are related by $S(d) = \prod_{i=1}^{d-1} (1 - \lambda(i))$.

Since $D_i$ is a discrete random variable, $f_D(d) \geq 0$ for all $d$ and $\sum_{i=1}^{\dmax} f_D(i) = 1$.
This implies that $F_D(d)$ is monotonically increasing.
Therefore, $S(d)$ is monotonically decreasing.
Both $F_D$ and $S$ are in the interval $[0, 1]$.
$\lambda(d)$ is in the interval $[0, 1]$, and can otherwise vary arbitrarily (\ie there is no monotonic or sum constraint).

\section{Infection process} \label{inc-prev:sec:infection-process}
The infection process is the mathematical description of incidence.
It is a \emph{point} or \emph{counting} process, where each infection is an event.

@@

Feedback loops mean that stochastically low or high numbers of infections affect the future number of infections.
That is, $\nu(t+\Delta t)$ is dependent on $K(t)$, even after conditioning on $\nu(t)$, for $\Delta t > 0$.
This violates the assumption of independent increments.
The reasons infectious diseases generate feedback loops is because stochastically high or low incidence feeds back to the process intensity at future times.
This can be seen most obviously in the event of a super-spreading event.
If a super-spreading event occurs, then many more individuals than suggested by the process intensity will be infected.
These additional infections will propagate the epidemic further, increasing the intensity at future times.
Point processes can be extended to \emph{branching processes} to account for this feedback~\autocite[246]{yanDistribution}, although these are beyond the scope of this thesis.

>>>>

Duration, or time-to-event, distributions can be expressed in several equivalent ways.
The duration distribution I will be concerned with is the distribution of the number of days for which an individual is prevalent.
Consider an individual $i$ infected on day $B_i$ that is prevalent during the interval $[B_i, E_i]$.
I define this period as an \emph{infection episode}, beginning on day $B_i$ and ending on day $E_i$.
$D_i = E_i - B_i + 1$ is their \emph{duration of positivity}, the number of days for which they are prevalent.
I model each $D_i$ as a discrete independent and identically distributed (iid) random variable; extending the methodology to account for covariates is left to further work.
The duration is a positive integer, \ie $D_i \in \{1, 2, \dots, \dmax \}$, where $\dmax$ is the maximum duration.
The first way of expressing the distribution of $D_i$ is as the standard probability mass function (pmf), $f_{D}(d) = \prob(D = d)$; or cumulative density function (CDF), $F_{D}(d) = \prob(D \leq d) = \sum_{k=1}^d f_{D}(k)$.
The \emph{survival function} gives the probability that the duration is at least $d$ days: $S(d) = \prob(D \geq d) = 1 - F_{D}(D - 1)$ with $S(1) = 1$ (a common alternative convention for survival analysis with discrete variables is to define the survival function as $\prob(D >d)$).
The \emph{hazard} on day $d$ is the probability of recovering at time $d$ conditional on still being positive on day $d$: $\lambda(d) = \prob(D = d \mid D \geq d) = f_{D}(d) / S(d)$.
Throughout this thesis, the survival function will be the primary quantity of interest, and it will normally be parameterised in terms of the hazard; these two quantities are related by $S(d) = \prod_{i=1}^{d-1} (1 - \lambda(i))$.

Since $D_i$ is a discrete random variable, $f_D(d) \geq 0$ for all $d$ and $\sum_{k=1}^{\dmax} f_D(k) = 1$.
This implies that $F_D(d)$ is monotonically increasing.
Therefore, $S(d)$ is monotonically decreasing.
Both $F_D$ and $S$ are in the interval $[0, 1]$.
$\lambda(d)$ is in the interval $[0, 1]$, and can otherwise vary arbitrarily (\ie there is no monotonic or sum constraint).

\section{Infection process} \label{inc-prev:sec:infection-process}
The infection process is the mathematical description of how infections occur.
It is a \emph{point} or \emph{counting} process, where each infection is an event.

@@

Feedback loops mean that stochastically low or high numbers of infections affect the future number of infections.
That is, $\nu(t+\Delta t)$ is dependent on $K(t)$, even after conditioning on $\nu(t)$, for $\Delta t > 0$.
This violates the assumption of independent increments.
The reasons infectious diseases generate feedback loops is because stochastically high or low incidence feeds back to the process intensity at future times.
This can be seen most obviously in the event of a super-spreading event.
If a super-spreading event occurs, then many more individuals than suggested by the process intensity will be infected.
These additional infections will propagate the epidemic further, increasing the intensity at future times.
Point processes can be generalised to \emph{branching processes} to account for this feedback~\autocite[246]{yanDistribution}, although these are beyond the scope of this thesis.
